\section{Datacentres}

CERN's computing equipment is located in two different datacentres: the
main one in Meyrin (Switzerland) and the secondary one in Wigner
(Hungary).

All the computing services provided by CERN like email, web hosting,
scientific processing and videoconferencing use the equipment based in the
datacentres. Both the datacentres are monitored 24/7/365 by on-site
technicians.

The datacenter in Meyrin is the oldest and biggest one: all the data
produced by the LHC is stored here and then distributed though the grid to
other centers around the world. In total there are more than 130 Petabytes
of stored data coming from the experiments and, considering that the LHC
produces around 30 Petabytes a year, this number expected to grow in the
following years. Long-term storage uses tapes to save huge amount of data
in a cost-efficient way: even if the technology is quite old, tapes are
able to store a large amount of data with smaller costs and in a small
space. Many tape robots ensure efficient storage and retrieval of data.

The datacentre in Wigner was inaugurated in June 2013 as an extention of
the Meyrin's one, it provides the extra computing power required to cover
CERN's increasing needs. All the equipment is managed and operated from
CERN, only the activities requiring physical access to the servers are
performed by the on-site technicians. Right now the secondary datacentre
has around 30\% of the capacity of the Meyrin's one, the aim is to have
the same capacity as the main one in 2020. The two datacentres are
connected using two independent and dedicated 100 Gbit/s fiber optic
lines, that result in approximately 25 microseconds of latency between
machines that are 1800km apart.

Some statistics about the datacentres:

% TODO: Update data
\begin{table}[H]
    \begin{center}
    \begin{tabular}{|l|c|c}
        \hline
        Number of Cores in Meyrin & 111,203\\
        \hline
        Number of Drives in Meyrin & 68,809\\
        \hline
        Number of Memory Modules in Meyrin & 75,263\\
        \hline
        Number of 10G NIC in Meyrin & 4,457\\
        \hline
        Number of 1G NIC in Meyrin & 21,013\\
        \hline
        Number of Processors in Meyrin & 20,259\\
        \hline
        Number of Servers in Meyrin & 10,961\\
        \hline
        Total Disk Space in Meyrin (TB) & 118,122\\
        \hline
        Total Memory Capacity in Meyrin (TB) & 434\\
        \hline
    \end{tabular}
    \caption[]{Datacentre Meyrin}
    \vspace{2em}
    \end{center}
\end{table}
\begin{table}[H]
    \begin{center}
    \begin{tabular}{|l|c|c}
        \hline
        Number of Cores in Wigner & 43,248\\
        \hline
        Number of Drives in Wigner & 23,169\\
        \hline
        Number of Memory Modules in Wigner & 21,598\\
        \hline
        Number of 10G NIC in Wigner & 1,399\\
        \hline
        Numer of 1G NIC in Wigner & 5,058\\
        \hline
        Number of Processors in Wigner & 5,408\\
        \hline
        Number of Servers in Wigner & 2,707\\
        \hline
        Total Disk Space in Wigner (TB) & 71,718\\
        \hline
        Total Memory Capacity in Wigner (TB) & 172\\
        \hline
    \end{tabular}
    \caption[]{Datacentre Wigner}
    \vspace{2em}
    \end{center}
\end{table}
\begin{table}[H]
    \begin{center}
    \begin{tabular}{|l|c|c}
        \hline
        Tape Drives & 104\\
        \hline
        Tape Cartridges & 26,463\\
        \hline
        Data Volume on Tape (TB) & 113,543\\
        \hline
        Free Space on Tape (TB) & 51,192\\
        \hline
        Routers (GPN) & 135\\
        \hline
        Routers (TN) & 29\\
        \hline
        Routers (Others) & 96\\
        \hline
        Star Points & 643\\
        \hline
        Switches & 3,533\\
        \hline
    \end{tabular}
    \caption[]{Network and Storage}
    \end{center}
\end{table}

