\section{Packaging with Koji}

At CERN we use Koji to build and ship packages to repositories and,
ultimately, to all the servers. Koji is a tool used by the official Fedora
Project \cite{FedoraProject} for packaging software and organize the
packages in different repositories. It is based on the concept of chroot
environments: every package is build in a new, clean environment in order
to avoid possible contaminations or ship unnecessary files.

Koji represents a main component of our infrastructure since it is
responsible for building the software before shipping it to all the
machines. A user is able to define a \textit{spec file} which is
a configuration file containing the instructions that Koji uses to build
the package.

In the \textit{spec file} we define all the instructions to build our
package correctly: the package name, version, release, architecture,
installation instructions, etc.

Our \textit{spec file} is able to generate two different packages from the
same repository, this allows us to build at the same time the reporter and
the cli for PackageInventory and deploy them together. Inside the
\textit{spec file} we specify all the files that are necessary to execute
PackageInventory correctly.

Specifying dependencies between packages (in our case, python libraries)
or configuration files is extremely easy. Moreover is fundamental to
specify configuration files inside the \textit{spec file}, due to the fact
that they do not need to be replaced in case of a package upgrade.

Developers can build their packages using Koji and then assign them to the
appropriate repositories. Koji takes care of building the package
correctly, following the specifications of the \textit{spec file}, for
different architectures.

The developer is allowed to build the package using Koji as many time as
it wants before actually publishing the package to the final repository.
This is very useful when we would like to test a build: we can trigger
a scratch build on Koji, download the resulting package and test it in our
test environment. In this way we can be sure that our software has been
packaged correctly before deploying it in the production servers.

Once the package is ready to be deployed it has to follow the usual
Configuration Change Procedure (CRM). The package is deployed for the
first week in a QA repository; all the QA machines point to this
repository and they will receive the update at the next Yum run. After one
week the package can be moved to the stable repository and then be
deployed to all the machines.

%TODO - Add code examples
