
\thispagestyle{empty}

Today, a three year long trip comes to conclusion.

It has been a hard one, with lot of ups and downs. It is useless to say
that I would not have made it without the support of all the people that
helped me during this amazing journey and made it as special as it was.

First of all, I want to thank my parents and my family for supporting my
studies and my education. They have been of great support and help during
all those years. Thank you for always backing my ideas and trusting me, it
is great to know that you believe in me.

After that, I would like to acknowledge all the Configuration Team at
CERN: Ben, Alberto, Nacho, Akos, Jerome, Ulrich, Gavin, Steve, Ignacio,
Fernando, Lorena, Iain and Paolo. They supported my projects and ideas all
the time, giving me opportunity to be really part of the team and not
a simple student.

I would like to thank my professor, Anna Ciampolini, for helping me
writing this thesis. Her suggestions have been extremely useful to me.

A huge thanks goes also to the house of Emilia Noyeres: to Adrian,
Radoslav, Aimilianos, Bruno, Adrian, Cristina and Niki. My time in Geneva
would not have been the same without the group of friends that you have
been for me. Thank you for all the moments we spent together, I will
never forget them.

Thank you to my oldest friends: Paolo, Nicola, Luca, Lorenzo, Giacomo and
Giulio. Even if we do not manage to see each other often, when we meet is
great for me to always find back the friends of a lifetime.

Thanks to all the people that I met during this amazing journey: you made
my life special, filling it with new experiences and a strong friendship.

Last, but not least, thanks to Fanny for her support during this last
year. Your words helped me when I thought that I could not make it. It has
been a long year for us, I am glad we made it through stronger than
ever.

Now new adventures are waiting and it is time to leave this university
that has seen me growing for many years. Changing habits and getting out
of the comfort zone is always scary at the beginning, but I believe it is
the only way to grow as a person. 

Anyway, the rule is always the same: don't panic.\\

\begin{center} So long, and thanks for all the fish \\ \end{center}

\begin{flushright} Andrea \end{flushright}
